\subsubsection{Journal Articles}
Takayuki Higo and Keiki Takadama.
Maintaining Multiple Populations with Different Diversities for Evolutionary
Optimization Based on Probability Models.
{\it IPSJ Transaction on Mathematical Modeling and Its Applications},
2008, to appear.
\\

\subsubsection{Conference Proceedings}
Takayuki Higo and Keiki Takadama.
Resampling-based Population Mechanism for Evolutionary Algorithms based on
Probability Models.
{\it 11th Asia-Pacific Workshop on Intelligent and Evolutionary Systems},
2007.
\\
\\
Takayuki Higo and Keiki Takadama.
Hierarchical Importance Sampling Instead of Annealing.
{\it IEEE Congress on Evolutionary Computation}, 2007.
\\
\\
Takayuki Higo and Keiki Takadama.
Neighbor based Parents Selection for Real-coded Genetic Algorithms
{\it Joint 3rd International Conference on Soft Computing and Intelligent Systems and 7th
International Symposium on advanced Intelligent Systems}, 2006.
\\
\\
Takayuki Higo and Keiki Takadama.
Comparison between Self-organization with Sampling and Genetic Algorithms in
multi-modal function
{\it International Symposium on Artificial Life and Robotics}, 2006.
\\

\subsubsection{Technical Reports and Conference Papers in Japanese}
比護貴之, 高玉圭樹.
多点探索型の最適化計算における収束問題のための階層型インポータンスサンプリング.
情報論的学習理論ワークショップ, 2006.
\\
\\
比護貴之, 高玉圭樹.
自己組織化サンプリングの提案とその最適化手法 -多峰性関数への適用-.
人工知能学会全国大会（第１９回）, 2005.
\\
\\
比護貴之, 高玉圭樹.
高次元多峰性関数最適化手法の提案.
計測自動制御学会　システム・情報部門 システム・情報部門学術講演会, 2005.
\\
\\
比護貴之, 高玉圭樹.
適応的探索領域分割に基づく最適化手法の提案.
計測自動制御学会　システム・情報部門 システム・情報部門学術講演会, 2004.
\\
\\
